---
title: "Coursework 3"
author: '36071280'
date: "02/11/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE,  message= FALSE}
library(tidyverse)

knitr::knit_hooks$set(timeit = local({
  now = NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res = difftime(Sys.time(), now)
      now <<- NULL
      # use options$label if you want the chunk label as well
      paste('Time for this code chunk:', as.character(res))
    }
  }})
)
```

# Question 1
Q1. Read in the data and print the dimensions of the Storm Events data frame

```{r readdata, warning=FALSE}
storm_events <- read.csv("Australia_severe_storms_1975-2015.csv")
print(dim(storm_events))

```

# Question 2
Q2. Clean the data by removing the variable ID and also Waterspout events from the database. Print
the dimensions of the cleaned data frame. Also print the first few rows without the 6 columns
of comments, without creating an intermediate data frame.

```{r cleandata}

# Removing variable ID and Waterspout event
storm_events_clean <- storm_events %>% select(-(ID)) %>% 
  filter(storm_events$Database != "Waterspout")
print(dim(storm_events_clean))
```

# Question 3
Q3. Add a column to your data frame containing the time zone of each event using the following OlsonNames() classifications. 

```{r Declaring Australian timezones and varibles}

#declaring variables for timezone mapping

list_of_relevant_australian_tz <- c("QLD" = "Australia/Queensland",
                                    "NSW" = "Australia/NSW",
                            "NSW_BrokenHill" = "Australia/Broken_Hill",
                            "VIC" = "Australia/Victoria",
                            "SA"="Australia/South",
                            "WA" = "Australia/West",
                            "TAS"= "Australia/Tasmania",
                            "NT" = "Australia/North",
                            "ACT" = "Australia/ACT")

new_south_wales = "NSW"
aus_central_time = "ACT"
broken_hill = "NSW_BrokenHill"
broken_hill_expr = "broken hill"
```

```{r Timezone allocation function}

allocate_tz <-function(state, nearest_town) {
  
  if (length(state) && !is.na(state))
  {
    if (state != new_south_wales)
    return (list_of_relevant_australian_tz[state])
    else if (state == new_south_wales)
    {
    
    if (length(nearest_town) && !is.na(nearest_town) && str_detect(tolower(nearest_town), broken_hill_expr))
      return (list_of_relevant_australian_tz[broken_hill])
    else
      return (list_of_relevant_australian_tz[new_south_wales])
    }
  }
  return (list_of_relevant_australian_tz[aus_central_time])
}
```

```{r Allocating correct timezone, time_it = TRUE}
storm_events_tz <- storm_events_clean %>% mutate(AustralianTimeZone = NA)

for ( i in 1:nrow(storm_events_clean))
{
  storm_events_tz$AustralianTimeZone[i] <- allocate_tz(storm_events_clean$State[i], storm_events_clean$`Nearest town`[i])
}

# storm_events_tz2 <- storm_events_clean %>% mutate(AustralianTimeZone = NA)  
# storm_events_tz2$AustralianTimeZone <- sapply(storm_events_tz2$AustralianTimeZone,
#                                               FUN = allocate_tz(storm_events_tz2, nearest_town = storm_events_tz2$Nearest.town))

```

# Question 4
Q4. Parse the date, time and time zones from the necessary columns to create a new variable in the data
frame which converts the time into UTC. You may need the function lubridate::as_datetime()
and/or the use of loops. Print the first few rows of the resultant data frame, without the 6 columns
of comments, again without creating an intermediate data frame.

```{r UTCdate}
library(lubridate)

storm_events_utc = data.frame()
for ( s in  list_of_relevant_australian_tz)
{
  storm_events_utc <- rbind(storm_events_utc, storm_events_tz %>% 
                              filter(storm_events_tz$AustralianTimeZone == s) %>%
                              mutate(UTCDateTime = (Date.Time %>% dmy_hm(tz = s) 
                                                    %>%  as_datetime())))
}

storm_events_utc %>% arrange(storm_events_utc$Event.ID)
```


```{r UTCdate For loop, eval =FALSE}

storm_events_utc2 <- storm_events_tz %>% mutate(UTCDateTime = NA)

for ( j in 1: nrow(storm_events_utc2)) {
  storm_events_utc2$UTCDateTime[j] <- (dmy_hm(storm_events_utc2$Date.Time[j], tz=storm_events_utc2$AustralianTimeZone[j]))
}
storm_events_utc2$UTCDateTime <- storm_events_utc2$UTCDateTime %>% as_datetime()

#comment_columns = c(...10,...11, ...12,...13, ...14, Comments)
# filter all the tz by the name,
# mass apply dmy_hm  -> loop should be 1:10
#head(storm_events_utc2 %>% select(-))
```


#Q5. 5. Create new variables for the month and year of each event. Print the first few rows of the resultant data frame, without the 6 columns of comments and without creating an intermediate data frame.

```{r months and years}

storm_events_with_cols = storm_events_utc %>% mutate(month_storm = month(storm_events_utc$UTCDateTime), 
                                                     year_storm = year(storm_events_utc$UTCDateTime))

#...10,...11, ...12,...13, ...14, Comments
print(storm_events_with_cols[-c(8:13)])
```

# Q6. After discarding Waterspout events there are five types of events left in the data; Rain, Hail, Lighting, Wind, and Tornado.

##i) Create a new data frame which contains the total number of counts for each of the above type of events for each of the twelve months over the forty year period. 

```{r}

storm_cols_grouped = storm_events_with_cols %>% group_by(Database, month_storm)
storm_counts_df = data.frame( count(storm_cols_grouped))
```

## ii)On a single plot, plot the total number of counts of each event against month. Use the R object month.abb for the labels of the months in the plot.

```{r plotting}
storm_counts_df$month_storm <- factor(storm_counts_df$month_storm, labels=month.abb)

ggplot(storm_counts_df,aes(factor(Database), n, fill=month_storm )) + 
  geom_bar(position="dodge2", stat="identity")

#ggplot(storm_counts_df,aes(factor(Database), n, fill=month_storm )) + 
 # geom_line()
```


# Q7. From the answer to Question 5, the 6 columns titled Comments, X, X.1, X.2, X.3, X.4 consist of comments.

## i) Combine the comments from these columns into a single column, named All.comments.


```{r}

storm_events_with_cols[is.na(storm_events_with_cols)] <- ''
storm_counts_all_comments <- storm_events_with_cols %>% mutate(AllComments = str_trim(paste(storm_events_with_cols$...10, storm_events_with_cols$...11,
                                                                                   storm_events_with_cols$...12, storm_events_with_cols$...13,
                                                                                   storm_events_with_cols$...14, storm_events_with_cols$Comments)))

#glimpse(storm_counts_all_comments)
```

## ii) Select the following columns to keep for further analysis, Event.ID, Database, State, All.comments, and the year variable you created.

```{r}
#storm_counts_all_comments$`Event ID`
#select(storm_counts_all_comments, storm_counts_all_comments$`Event ID`)
storm_analysis <-  storm_counts_all_comments %>% select(c(Event.ID, Database, State, AllComments, year_storm))

print(sapply(storm_analysis, class))
```

# Q8. Now we use the answer to Question 7(ii) for further analysis

## i) Create an indicator variable which states whether or not a storm event has resulted in a flash flood. Make sure you sort out all terms relating to flash floods.

```{r}
flash_flood_expr <- "flash.*?flood*" # flash - any? -  flood 
flash_flood_expr_space_mul <- "flash\\s*?flood*" # flash - space(s)? -  flood 
flash_flood_expr_mul <- "flash(.|\\s*)?flood*" #final

storm_analysis <- storm_analysis %>% mutate(flash_flood_present = str_detect(storm_analysis$AllComments,                                                                             regex(flash_flood_expr_mul, ignore_case = TRUE)))

storm_analysis_with_floods <- storm_analysis %>% filter(flash_flood_present==TRUE)
storm_analysis_grouped <- storm_analysis_with_floods %>% group_by(year_storm)

dim(storm_analysis_grouped)

```

## ii) Print a plot of the number of flash floods per year from 1975-2015. You may need to first create a vector or data frame to contain the number of flash floods per year.

```{r}

flash_flood_counts <- count(storm_analysis_grouped)

ggplot(flash_flood_counts) + geom_line(aes(year_storm, n))
ggplot(flash_flood_counts,aes(year_storm, n)) + geom_bar(stat="identity")

#View(flash_flood_counts)
```


# Q9. For severe wind events often the wind speed is given. The wind speed is given in knots or km/h.

## i) Extract all wind speeds both those in knots and km/h. Hint: Knots can be abbreviated by kts or kt. Also note that wind speed can be a single, double or triple digit number

```{r}
wind_speed_expr <- "(\\d{1,3})\\s?(knot|kt(s)?|km\\s?\\/\\s?h(r)?)"

storm_analysis_speeds <- str_extract_all(storm_analysis,regex(wind_speed_expr))

storm_analysis_speeds <- storm_analysis %>% mutate(wind_speed = str_extract( storm_analysis$AllComments,wind_speed_expr))

storm_analysis_speeds <- storm_analysis_speeds %>% filter(!is.na(wind_speed))

#write.csv(storm_analysis_speeds, "storm_analysis_speeds.csv")
```

## ii) Convert km/h wind speeds to knots (1 knot = 1.852 km/h) rounding the wind/speed to the nearest knot. Hint: It is helpful to work with a reduced data frame which includes only those observations with a wind speed recorded.

```{r}
knot_to_kmh = 1.852
wind_speed_expr_km <- "(\\d{1,3})\\s?(km\\s?\\/\\s?h(r)?)"
extract_number_expr <- "\\d+"

get_wind_value <- function (wind_speed) {
  wind_numeric_value = as.numeric(str_extract(wind_speed, extract_number_expr))
  if (str_detect(wind_speed, regex(wind_speed_expr_km, ignore_case = TRUE)))
  {
    return (as.integer(wind_numeric_value/knot_to_kmh))
  }
  else
  {
    return (wind_numeric_value)
  }
}
```


```{r}

storm_analysis_speed_numbers <- storm_analysis_speeds %>% mutate(wind_values = NA)
storm_analysis_speed_numbers$wind_values <- sapply(storm_analysis_speed_numbers$wind_speed, FUN = get_wind_value)

#as.numeric(str_extract(storm_analysis_speeds$wind_speed[27], extract_number_expr))
#storm_analysis_speed_numbers[1:15,]
```

## iii) Print a boxplot of the wind speeds recorded per state. 

```{r}

ggplot(storm_analysis_speed_numbers, aes(State, wind_values)) + geom_boxplot(position="dodge2")
```

