---
title: "SCC 461 Coursework 0"
author: '36071280'
date: "10/10/2021"
output:
  pdf_document: default
  word_document: default
---

Q1. Write code for log(PCB) against age, reproducing the final plot containing both the equation line and
the data points.

```{r}
#Task 1 - Reproducing the plot for age vs. log(PCB)

ages <- seq(from=0, to=13, by=0.1)
a <- -2.3907
b <- 2.300
l <- a + b*ages^(1/3)
trout.age <- c(1, 1, 1, 1, 2, 2, 2, 3, 3, 3, 4,4, 4, 5, 6, 6, 6, 7, 7, 7, 8, 8,
               8, 9, 11, 12, 12, 12)
trout.pcb <- c(0.6 , 1.6, 0.5, 1.2, 2.0, 1.3, 2.5,2.2, 2.4, 1.2, 3.5, 4.1, 5.1, 5.7,
               3.4, 9.7, 8.6, 4.0, 5.5, 10.5, 17.5,13.4, 4.5, 30.4, 12.4, 13.4, 26.2, 7.4)
plot(x=trout.age, y=log(trout.pcb), main="Log(PCB) vs Age", xlab = "Age ->", ylab = "Log(PCB) ->")
lines(x=ages, y=l, type="l", col="blue")

```

Q2. Rewrite the log(PCB) equation as a function which has arguments; a, b, and age, and returns the
predicted log(PCB).

``` {r}

#Task 2 - Rewrite log(PCB) as a function

calc_PCB = function(a, b, age) {
  predicted_l <- a + b*age^(1/3)
  return (predicted_l)
}

```

Q3. By extending the range of age considered, produce a plot which shows the curve for the expected
log(PCB) concentration for lake trout up to 20 years old.


``` {r}

#Task 3 - Extend ages to 20 years

ages_20 <- seq(from=0, to=20, by=0.1)
plot(x=ages_20, y=calc_PCB(a,b,ages_20), "l", col="green", main="Log(PCB) vs Age for 20 years", 
     xlab="Age ->", ylab="Log(PCB) ->")
```

Q4. Now extract the maximum expected/predicted log(PCB) from the values used to draw the equation
line

``` {r}
#Task 4 - Get Maximum
print(paste0("Maximum value is :", max(calc_PCB(a,b,ages_20))))
```
Q5. It can be shown that a non-linear model of the form
l = a + b × age^c
where a, b, and c are constants provides a slightly better fit to the data. The optimal choices are a = -4.865, b = 4.7016, and c = 0.1969.

```{r}

#Task 5 - Non-linear model

a2 = -4.865
b2 = 4.7016
c2 = 0.1969

```

(a) Rewrite the log(PCB) equation as a function which has arguments; a, b, c and age, and returns the
predicted log(PCB).

``` {r}

# Task 5.1 - Rewrite log equation function

calc_PCB2 = function(a,b,c,age) {
  predicted_l = a + b * (age^c) 
  return (predicted_l)
}

```

(b) Compare the Bates-Watts estimator and the new estimator for the expected log(PCB) concentration of
a 10 year old lake trout.

```{r}

# Task 5.2 - Compare

fixed_age = 10
value_BWEstimator = calc_PCB(a,b,fixed_age)
value_newEstimator = calc_PCB2(a2,b2,c2,fixed_age)

value_newEstimator == value_BWEstimator
```

(c) Create a new plot which has both the old line and new line, allowing a comparison of the differences. 


```{r}

# Task 5.3 - Plot both equations

plot(x=ages, y=l, type="l", col="red", ylab = "Log(PCB) ->", xlab = "Ages ->", 
     main="Comparing Bates-Watts estimator and the new estimator")
lines(ages, calc_PCB2(a2,b2,c2,ages), type="l", col="orange")

```